---
title: "HCR probe design"
format: html
editor_options: 
  chunk_output_type: console
---

```{bash}
<!-- conda create -n BLAST python=3.7 -->
<!-- conda activate BLAST -->
<!-- conda install -c bioconda blast -->

<!-- cd /PATH/TO/FASTA/File -->

<!-- makeblastdb -in Dmel_ens99_from-MK.fa -parse_seqids -title Dmel_ens99 -dbtype nucl -->

```


```{r}
# Sys.setenv(PATH = paste(Sys.getenv("PATH"), "/opt/miniconda3/envs/BLAST/bin", sep = .Platform$path.sep))
# Sys.getenv("PATH")

```


## Environment

```{r}
library(tidyverse) # CRAN
library(furrr) # CRAN
library(patchwork) # CRAN
library(valr) # CRAN
library(Biostrings) # Bioconductor
library(rBLAST) # Github

Sys.which("blastn") # Check correct installation of BLAST
n_threads <- 4
plan(multisession, workers = n_threads) # Number of threads
source("./src/HCR_design.R") # Contains functions

```

## Prepare target RNA sequence

Put N between exons to prevent probes spanning exon-intron junctions

```{r}
## Paste target RNA sequence
target_raw <- "ATTCACATCGGTGCAAGCGTTCCGCAATCTGCTCAAGGAGGTGGGCGATC
TGGCGGGACAGCGCGAGGTGGTGTCCGAGTCCCTGCAGCTGCAGATCATT
GCGGGAGTGACGCTTCTGTCCAAGACATTGCGCGAGGAACGCAAG N 
AAATGCCTTAGCGATGGTGCCACCCTGCAGCAGAACCTCACCACACAGCT
CTCCTCGCTGGACCGGGCCAAGCGGAACTACGAGAAGGCCTACCGTGACT
CGGAGAAGGCGGTGGACAGCTATAAGCGGGCAGACATGGACCTCAATCTC
AGCCGGGCCGAGGTGGAGCGCTACAAGAACGTGATGACGTCCAAGATCCA
GCAGTCGGACGATGCGAAGAACGAGTACGCTAACCAGCTACAGAAGACGA
ACAATCTGCAGCAGCAACACTACAGCATGCTGCTGCCCTCGGTCCTCAAT
CGGCTGCAGGAGCTGGACGAGAAACGCACCCGTGGCTTCAGGGAGTTCAT
TGTGGGAGCGGCGGATGTGGAGTCATCGGTGGCGCCAATCATAGCCCGCT
GCATGGAGGGTATCGTGAAGGCCGGCGAGTCCATCAACGAAAAGGAGGAT
ACCTTCAAAGTCATAGAAAG N 
ATATCAATCTGGTTTCACGCCACCAAGGGACATACCCTTCGAGGATCTGT
CCAAGTGCGATCCGGATTCCGTGCAGGACTCACACTACAGCAACTCGACA
TCGAACCACCTGACCATTAGAGGCACGATGAGTGCCAACAAGCTGAAGAA
ACGCGTGGGCATTTTTAACATATTCGGCAGCAATAAG N
CAGTTCGGCAAGTCCCGAGAGTGGCCTTGGCACTTCGCACACATCCCTGC
CAGGATCAGGACAGGGCAGCGCCAACGAAAACGCGATTGGCGAGGATACC
TACTATGAAACGGAAGTGGAGACCCTTAATCCAGTGGGAAAATGTCGTGC
CCTGTATCCATTTGAAG N
CCTCCAGCGAGGGTAGCATACCCATGAGTGAGGGCGAGGAGCTGCAAGTG
ATCGAGATCGACCAAGGAGACGGATGGACGCGGGTGCGGCGGGAGAACAA
CTCCAATGGCTGGGACGAGGGCTTCGTGCCCACGAGTTACATCGAGATCA
CGCTCTATGCTTAG N
GATTAAGTTCATTTTCGGGGCGGGCAATCAGCCGGGCGGTTTTAAATGTT
ACCTACTCAGCAAACTCTTGCGTATGATCGATTGATTTGTGTACTTTTGT
CATAACTTTGCGTTCTGTCGGACAGCCATAACACACAAAACATAACATAA
CAGACACAACACGCCTAACTTTCACACAACACTGTACATTTGACTTAAGC
TAAGTTCGATTCATTTACCCTTCGTATGCTTTGTATATTTACTTTGATTT
TAAACGTCATTGATCTTATGGTAGTGCAAGCTATATCATACTATAAATAC
TATAGTCATATGCACCTACGAGCAGACCTCACCTAATTGTAAGCACCGCA
ACGTTTCCCTCTTCACCACAACAATCATGATAACACCTATTTATATACAT
ATATGGAAAAAGATTTATATGACACTTGAATACGTGAGAGCATGAAAAGG
AAAACTTAAACTGAAACTTAAACTACAGCCGTTGGACATTAAAAGTAATC
TTGTGAACACGTTCTCACGTGCAATCC
"
fasta_file <- "./input/hTMPRSS2-test.fa" # Each exon should be separate fasta records
target_fasta <- readDNAStringSet(fasta_file)
target_raw <- paste(target_fasta, collapse = "N") # Alternatively, paste sequence into target_raw directly

## Prepare target RNA sequence
target <- clean_pasted_seq(target_raw)
target_sequence_total_length <- nchar(target)

```

## Extract candidate probes into a dataframe

```{r}
## Extract candidate probes into a dataframe
target_name <- "hTMPRSS2-test"
oligo_length <- 52
candidate_probes <- generate_candidate_probes(target_seq   = target,
                                              oligo_length = oligo_length)
## Inspect
inspect_probe_prep(candidate_probes)

```

## Annotate candidate probes for filtering metrics

hairpin_dG, homodimer_dG 

```{r}
## Get thermodynamic parameters - RNA sequence as input!! (Using furrr package to speed up processing)
candidate_probe_thermodynamics <- future_map(
  candidate_probes$target_sequence, 
  ~ calculate_thermodynamics(seq           = .x, 
                             temperature   = 37,
                             Na            = 0.3,
                             oligo_conc    = 5e-5)
)

candidate_probe_thermodynamics_1st_half <- future_map(
  map_chr(candidate_probes$target_sequence, ~ str_sub(.x, start = 1, end = nchar(.x)/2 - 1)),
  ~ calculate_thermodynamics(seq         = .x,
                             temperature = 37,
                             Na          = 0.3,
                             oligo_conc  = 5e-5)
)

candidate_probe_thermodynamics_2nd_half <- future_map(
  map_chr(candidate_probes$target_sequence, ~ str_sub(.x, start = nchar(.x)/2 + 2, end = nchar(.x))),
  ~ calculate_thermodynamics(seq         = .x,
                             temperature = 37,
                             Na          = 0.3,
                             oligo_conc  = 5e-5)
)


## Annotate candidate sequences 
candidate_probes_annotated <- annotate_probes(candidate_probes)

```

## Filter based on thermodynamics / base composition parameters 

- Raj lab HCR (52nt)
 - dG [-70, -50] https://journals.asm.org/doi/10.1128/mbio.03751-21 
- Sakura-Spengler lab (25nt) 
 - GC [37, 85]
 - Tm [55, 77]
- Oligostan 26-32nt
 - dG -32
 - GC [40, 60]
 - Nucleotide composition rules

```{r}
## Generate exploratory plots
c("Tm", "dG", "GC_content") %>% generate_exploratory_plots()
```

```{r}
## Filter based on thermodynamic/nucleotide compositions
target_dG <- -60
target_dG_halves <- -32
dG_range <- c(-75, -50)
Tm_range <- c(10, 100)
GC_range <- c(0.4, 0.7)
pass_a_comp       <- FALSE
pass_c_comp       <- FALSE
pass_a_stack      <- TRUE
pass_c_stack      <- TRUE
pass_c_spec_stack <- TRUE

candidate_probes_filtered <- filter_candidate_probes(candidate_probes_annotated)

## Inspect filtered probes
plot_inspection(candidate_probes_filtered, "Tm")

```


## BLAST screen

BLAST screen with short blastn mode.
RMBLAST or Bowtie1? Instead of bowtie2?

```{r}
## Params
blast_file <- "~/Documents/BLAST/Dmel/Dmel_ens99_from-MK.fa"
tx2gene_csv <- "./src/Dmel_tx2gene_ENSEMBL_v99.csv" # To convert tx_id to gene_id
max_blast_matches <- 1 # Number of BLAST hits
allowOverlapBreakRegion <- TRUE # Allow overlap over the 25/25 break region?

## Set up blast environment
db <- blast(db = blast_file)

## Get sequences of filtered candidate target_sequence 
seqs <- DNAStringSet(candidate_probes_filtered$target_sequence)
names(seqs) <- candidate_probes_filtered$unique_id

## BLAST nucleotide
blast_output <- run_blastn_short(db, seqs, tx2gene_csv)
blast_summary <- summarise_blast_output(blast_output, allowOverlapBreakRegion)

## Screen candidate probes
screened_ids <- blast_summary %>% 
  filter(n_matches <= max_blast_matches) %>% pull(qseqid)
candidate_probes_screened <- candidate_probes_filtered %>%
  filter(unique_id %in% screened_ids)

## Merge overlapping intervals and inspect  
candidate_probes_screened <- mutate(candidate_probes_screened, chrom = "valr")
merge_df <- bed_merge(candidate_probes_screened) %>%
  mutate(region_length = end - start + 1) %>%
  mutate(type = case_when(
    region_length >= 2* oligo_length ~ 1,
    region_length < 2 * oligo_length ~ 0,
    TRUE ~ 0
  ))

plot_inspection(candidate_probes_screened, "Tm") +
  geom_linerange(data = merge_df, 
      aes(x = "", ymin = start, ymax = end),
      colour = "gray40", size = 2,
      inherit.aes = FALSE)

```

## Distribute and select probes

```{r}
## Params
probe_spacing <- 3

## Get final non-overlapping screened candidate probes 
candidate_probes_final <- distribute_overlapping_probes(candidate_probes_screened, merge_df, probe_spacing = 3)

## Inspect
plot_inspection(candidate_probes_final, "Tm") +
  geom_linerange(data = merge_df, 
      aes(x = "", ymin = start, ymax = end),
      colour = "gray40", size = 2,
      inherit.aes = FALSE)

```

## Attach HCR-B initiator sequences 

Attach HCR B1/B2/B3/B4/B5 split initiator sequences to selected probe sequences (rev-com'ed)
 - Initiator b is added 3' end of First half (5') sequence
 - Initiator a is added 5' end of Second half (3') sequence

```{r}
# * * * * * CHOOSE HAIRPIN FOR MULTIPLEXING

b_identifier <- "B1"

# * * * * * Attach hairpin sequences and save 

probe_details <- attach_hcr_initiatior(candidate_probes_final, b_identifier = b_identifier)
probes <- get_final_probe_sequences(probe_details)

```

## Save probe sequences and some other outputs

```{r}
## Probe details
write_csv(probe_details,  paste0("./output/", target_name, "_", "HCR", b_identifier, "_details.csv"))

## Probe design parameters
save_params(paste0("./output/", target_name, "_", "HCR", b_identifier, "_params.txt"))

## Probe sequences
write_csv(probes, paste0("./output/", target_name, "_", "HCR", b_identifier, "_probes.csv"))

## Probe location plot
plot_final_probes(probe_details, "Tm")
ggsave(paste0("./output/", target_name, "_", "HCR", b_identifier, "_probes.pdf"), 
       width = 7, height = 4)
  
```












```{r}
install.packages("reticulate")

library(reticulate)

```


```{bash, engine.opts = "-l"}
conda activate BLAST
blastn -h

```






















/usr/people/bioc1402/data/computational_projects/C2/2.25.fishprobe_design/dmel_200729/offtarget_filtering/Dmel
[16:15] Mary Thompson
/usr/people/bioc1402/miniconda3/envs/ribopop_probe_design























































